---
title: "Kickstarter"
author: "Meredel Ababa"
date: "1/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
kickstarter = ks_projects_201801
```

```{r}
kickdata = ks_projects_201801 # import data
kickdata$state1 <- kickdata$state # made the state1 column 
kickdata$state1 <- gsub("canceled|undefined|live|suspended", "null", kickdata$state1) # replaced canceled, undefined, and live with null
kickdata <- kickdata[kickdata$state1 != "null", ]   # removed null
kickdata <- kickdata                 # code that removes scientific notation.
options(scipen=999)  # removes scientific notation
Kickdata 
kickdata$state1 = ifelse(kickdata$state =='failed', 0,1) # kickdata 0 is failed and 1 is successful
```

```{r}
summary(kickstarter)
```

```{r}

# creating launched1 as lubridate object
kickdata$launched1 =          lubridate::as_date(kickdata$launched)
# creating deadline1 as lubridate object
kickdata$deadline1 = lubridate::as_date(kickdata$deadline)
# checking to see all variables
names(kickdata)
# creating new duration column
kickdata$dayz <- kickdata$deadline1 - kickdata$launched1 # called dayz
head(kickdata)
# seeing summary on new data set w/new columns
kickdata$l_time <- format(as.POSIXct(kickdata$launched) ,format = "%H:%M:%S") #extracted time from launched column
kickdata$d_time <- format(as.POSIXct(kickdata$deadline) ,format = "%H:%M:%S") #extracted time from deadline column
launched_month = month(as.POSIXlt(kickdata$launched1, format="%Y/%m/%d")) #extract month from launched1 date
dead_month = month(as.POSIXlt(kickdata$deadline1, format="%Y/%m/%d")) #extract month from deadline1 date
kickdata=data.frame(kickdata, launched_month) #add launched month to df
kickdata=data.frame(kickdata, dead_month) #add deadline month to df
kickdata$launched_month=as.factor(kickdata$launched_month) # launched_month needs to be factor, so we convert factor
kickdata$dead_month=as.factor(kickdata$dead_month) #dead_month needs to be factor, so we convert to factor
```

```{r} # function that will take our date, transform to 2012 date, then find the season in which it occured
getSeason <- function(launched1) {
    WS <- as.Date("2012-12-15", format = "%Y-%m-%d") # Winter Solstice
    SE <- as.Date("2012-3-15",  format = "%Y-%m-%d") # Spring Equinox
    SS <- as.Date("2012-6-15",  format = "%Y-%m-%d") # Summer Solstice
    FE <- as.Date("2012-9-15",  format = "%Y-%m-%d") # Fall Equinox

    # Convert dates from any year to 2012 dates
    d <- as.Date(strftime(launched1, format="2012-%m-%d"))

    ifelse (d >= WS | d < SE, "Winter",
      ifelse (d >= SE & d < SS, "Spring",
        ifelse (d >= SS & d < FE, "Summer", "Fall")))
}

kickdata$launch2012 = as.Date(kickdata$launched1, format = "%Y-%m-%d") + 0:331674 #formatting to 2012 
kickdata$seasonlaunch = as.factor(getSeason(kickdata$launch2012)) #use function to find season for launch 
kickdata$launch2012 = NULL

kickdata$deadline2012 = as.Date(kickdata$deadline1, format = "%Y-%m-%d") + 0:331674 #formatting to 2012 
kickdata$seasondeadline = as.factor(getSeason(kickdata$deadline2012))  # use function to find season for deadline
kickdata$deadline2012 = NULL 
```

```{r}
#histogram of backers
hist(kickdata$backers [kickdata$backers])
```
#assessing what type of variable it is. 
```{r}
#seeing class of all our variables 
class(kickdata$state1)
class(kickdata$backers)
class(kickdata$goal)
class(kickdata$l_time)
class(kickdata$d_time)
class(kickdata$launched_month)
class(kickdata$dead_month)
class(kickdata$name)
class(kickdata$country)
class(kickdata$state)
class(kickdata$usd.pledged)
class(kickdata$pledged)
class(kickdata$launched1)
class(kickdata$dayz)
```

```{r}
# dropping variables that won't be included in analysis
kickdata2 = kickdata
kickdata2$ID = NULL
kickdata2$name= NULL
kickdata2$currency= NULL
kickdata2$deadline= NULL
kickdata2$launched= NULL
kickdata2$launched1= NULL
kickdata2$deadline1 = NULL
kickdata2$state = NULL
kickdata2$pledged = NULL
kickdata2$usd.pledged=NULL #too correlated with goal
kickdata$launchyear = format(as.POSIXct(kickdata$launched), format = "%Y") #extracting the correct format for launchyear 
```
changing to factor
```{r}
# changing variables to factor
kickdata2$state1 <- factor(kickdata2$state1) 
kickdata2$category <- factor(kickdata2$category)
kickdata2$main_category <- factor(kickdata2$main_category)
kickdata2$country <- factor(kickdata2$country)
kickdata2$l_time <- factor(kickdata2$l_time)
kickdata2$d_time <- factor(kickdata2$d_time)
kickdata2$launched_month <- factor(kickdata2$launched_month)
kickdata2$dead_month <- factor(kickdata2$dead_month)
kickdata2$dayz =as.numeric(paste(kickdata2$dayz))
kickdata2$backers =as.numeric(paste(kickdata2$backers))
```

split data to training and test set. Unfortunately, will not run.
```{r}
set.seed(123)
train <- sample(1:331675,331675*.70,rep=FALSE) # creating training set of data , 70%
test <- -train # test set, will be 30%, whatever is left from training set
kick.train <- kickdata2[train, ] # 
kick.test <- kickdata2[-train, ]
#log.regression = glm(state1~ goal+launched_month, data = kick.train, family = binomial(link="logit"))
#summary(log.regression)
```
Error rate is .437826
```{r}
qda_model = qda(state1~ launched_month+category+goal+seasonlaunch+dayz, data = kick.train) #QDA model 
qda_pred = predict(qda_model, kick.test)  # run the model on our test set
qda_pred_y = qda_pred$class
table(qda_pred_y, kick.test$state1)
mean(qda_pred_y != kick.test$state1)
```
#Better QDA model, but not by much. Error rate is .4370521
```{r}
library(MASS) 
qda_model = qda(state1~ category+ seasondeadline+goal+seasonlaunch+dayz, data = kick.train)
qda_pred = predict(qda_model, kick.test)
qda_pred_y = qda_pred$class
table(qda_pred_y, kick.test$state1)
mean(qda_pred_y != kick.test$state1)
```
#Performed even worse with error rate of .5602
```{r}
class(kickdata2$launched_month)
qda_model = qda(state1 ~ main_category + seasondeadline + goal + seasonlaunch+dayz+dead_month+launched_month, data = kick.train)
qda_pred = predict(qda_model, kick.test)
qda_pred_y = qda_pred$class
table(qda_pred_y, kick.test$state1)
mean(qda_pred_y != kick.test$state1)
```
Cross validation on QDA : error rate is 34.215, with a K at 10
```{r}
CV.qda <- #create CV function
  function (data, model, dname, K=10, seed=150) {
    n <- nrow(data)
    set.seed(seed)
    dataY=data[,dname]

    f <- ceiling(n/K)
    s <- sample(rep(1:K, f), n)

    CV=NULL

    for (i in 1:K) {
      test.out <- seq_len(n)[(s == i)] #test data
      train.in <- seq_len(n)[(s != i)] #training data


      qda.fit=lda(model, data=data[train.in,])
      qda.y <- data[test.out, dname]
      qda.predy=predict(qda.fit, data[test.out,])$class

      error= mean(qda.y!=qda.predy)
      CV=c(CV,error)
    }
    list(call = model, K = K,
         qda_error_rate = paste(100*mean(CV), "%"), seed = seed)
  }
err.qda=CV.qda(data = kickdata2, state1~main_category + seasondeadline + goal + seasonlaunch+dayz+dead_month+launched_month, dname = "state1", K=10, seed = 123) #checking error 
err.qda
```
#extracted year from launched so we can use that to compare projects by year
```{r}
kickdata$launchyear = format(as.POSIXct(kickdata$launched), format = "%Y")
```
```{r}
library(FNN)
kickdata3=kickdata2
kickdata3$category <- NULL
kickdata3$main_category <- NULL
kickdata3$country <- NULL
kickdata3$l_time <- NULL
kickdata3$d_time <- NULL
kickdata3$seasonlaunch <- NULL
kickdata3$seasondeadline <- NULL
kickdata3$X <- NULL
kickdata3$dayz =as.numeric(paste(kickdata3$dayz))
kickdata3$backers =as.numeric(paste(kickdata3$backers))
kickdata3$goal =as.numeric(paste(kickdata3$goal))
kickdata3$launched_month =as.numeric(paste(kickdata3$launched_month))
kickdata3$dead_month =as.numeric(paste(kickdata3$dead_month))
kickdata3$state1 =as.numeric(paste(kickdata3$state1))
```
```{r}
data = scale(kickdata3[,-c(3,4)]) # scale the variable; while also removing the predictor (state1)
set.seed(123) #setting seed
train <- sample(1:331675,331675*.7,rep=FALSE) # creating the training data set 70%
test <- -train # creating test data set
training_kick = data[train, ] 
testing_kick = data[test,]
summary(data)
```
#includes state variable
```{r}
training_y = kickdata3$state1[train] # training set
testing_y = kickdata3$state1[test] # testing set
```
library(class) #to be able to run kin()
set.seed(123) #setting the seed
knn_pred_y = knn(training_kick, testing_kick, training_y, k = 1) # KNN
#dim(training_data)
#length(training_y)
table(knn_pred_y, testing_y) #Provides the confusion matrix
mean(knn_pred_y != testing_y) #Misclassification error rate at 0.1133735
```
Created a CV KNN Function
```{r}
knn_pred_y = NULL
error_rate = NULL
# creating function for KNN and getting error rate
for(i in 1:50){
set.seed(123)
knn_pred_y = knn(training_kick,testing_kick,training_y,k=i)
error_rate[i] = mean(testing_y != knn_pred_y)
}

### find the minimum error rate
min_error_rate = min(error_rate)
print(min_error_rate)

### get the index of that error rate, which is the k
K = which(error_rate == min_error_rate)
print(K)
```
To visualize the misclassification error rate
```{r}
library(ggplot2)
qplot(1:300, error_rate, xlab = "K",
ylab = "Error Rate",
geom=c("point", "line"))
